---
title: "Task 4"
author: "Markus Bentsen"
format: html
editor: visual
---

```{r}
#Loading packages
rm(list=ls())
library(tidyverse)
library(lubridate)
library(rjstat)
library(janitor)
library(jsonlite)
library(gdata)
library(httr) 
library(ggthemes)
library(dplyr)
library(Quandl)
library(readr)
library(quantmod)

#URL 
df <- read_csv("https://raw.githubusercontent.com/uit-sok-1005-v23/uit-sok-1005-v23.github.io/main/storedata.csv")


```

# Task 1

1: For the last 3 months of 2017, calculate the total Sales by month, for Region 1 and Region 9 in the Customer_Segment, Corporate, and Consumer. This output is Table 1.

```{r}
#Task 1, segment 1

#Creating a variabel for month and year
df$Year <- format(df$Order_Date, "%Y")
df$Month <- format(df$Order_Date, "%m")

#Creating a dataframe for task 1
df_2017 <- df

#filtering out everything besides region 1 and 9
df_2017 <- df_2017 %>% 
  filter(Year == "2017", Month >= "10", Customer_Segment %in% c("Corporate", "Consumer"), Region %in% c("Region 1", "Region 9")) %>% 
  group_by(Region, Month, Customer_Segment) %>%
  summarize(Total_Sales = sum(Sales)) 


df_2017
```

2: Make a plot of the monthly total Sales in Region 1 and Region 13 in 2015, 2016, and 2017. This output is Figure 1.

```{r}
#Task 1, segment 2
figur1 <- df %>% 
  filter(Year %in% c("2015", "2016", "2017"), 
  Region %in% c("Region 1", "Region 13")) %>% 
  group_by(Region, Year, Month) %>%
  summarize(Total_Sales = sum(Sales)) %>%
  ggplot(aes(x = Month, y = Total_Sales, fill = Year)) +
  geom_col(width = 0.8, position = "dodge") +
  facet_wrap(~ Region, ncol = 2) +
  labs(x = "Months", y = "Sales", fill = "Year")+
  ggtitle("Total sales for region 1 and 13") +
  theme_bw()

figur1
```

3: In Figure 1, identify the months where the total Sales in Region 13 is greater than the total Sales in Region 1. This output is Table 2.

```{r}
#Filtering out every year besides 2015-2017 and region 1 and 13
table2 <- df %>%
  filter(Year %in% c("2015", "2016", "2017"), 
  Region %in% c("Region 1", "Region 13")) %>% 
  group_by(Region, Year, Month) %>%
  summarize(Total_Sales = sum(Sales))

#Comparing region 1 and 13 to see identify the months where total sales are higher
table2 <- table2 %>%
  pivot_wider(names_from = Region, values_from = Total_Sales) %>%
  filter(!!sym("Region 1") < !!sym("Region 13")) %>% 
  select(Year, Month, !!sym("Region 1"), !!sym("Region 13"))
 

table2
```

4: Find the average Profit per Customer_Segment and Product_Category in 2017, for all regions except Region 3, 5 and 8. What segment produced the highest average profit? This output is Table 3.

```{r}
#Filtering out everything besides 2017 and finding the mean for each customer
table3 <- df %>% 
  filter(Year == 2017, !Region %in% c("Region 3", "Region 5", "Region 8")) %>%
  group_by(Customer_Segment, Product_Category) %>% 
  summarize(Profitt = mean(Profit)) %>% 
  arrange(desc(Profitt)) 

table3
```

# Task 2

```{r}
#Downloading the required data from xom and making  it into a dataframe
getSymbols("XOM", src = "yahoo") 
xom_df <- as.data.frame(XOM)
xom_df <- tibble::rownames_to_column(xom_df, var = "Date")
xom_df$Date <- as.Date(xom_df$Date) 
xom_df <- xom_df %>% 
  filter(Date >= "2010-01-04" & Date <= "2022-12-31")

#getting the monthtly average
xom_monthly <- xom_df %>%  
  mutate(year_month = format(Date, "%Y-%m")) %>%
  group_by(year_month) %>% 
  summarize(exxon_avg = weighted.mean(XOM.Close, XOM.Volume))

#creating a date variabel
xom_monthly$year_month <- ym(xom_monthly$year_month)
```

```{r}
#Downloading data for brent oil and making it into a dataframe
getSymbols("DCOILBRENTEU", src = "FRED") 
oil_df <- as.data.frame(DCOILBRENTEU)
oil_df <- tibble::rownames_to_column(oil_df, var = "Date")
oil_df$Date <- as.Date(oil_df$Date)
oil_df <- oil_df %>%
  filter(Date >= "2010-01-04" & Date <= "2022-12-31")

#getting the montly average
oil_monthly <- na.omit(oil_df) %>%  
  mutate(year_month = format(Date, "%Y-%m")) %>% 
  group_by(year_month) %>% 
  summarize(oil_avg = mean(DCOILBRENTEU))

#Creating a date variabel
oil_monthly$year_month <- ym(oil_monthly$year_month)
```

```{r}
ggplot() + 
  geom_line(data = xom_monthly, aes(x = year_month,
  y = exxon_avg, color="XOM")) +
  geom_line(data = oil_monthly, aes(x = year_month, 
  y = oil_avg, color="OIL")) + 
  xlab("Year") +  
  ylab("Price") +
  ggtitle("Monthly average prices of Exxon and Brent oil") + 
  theme_bw() 
```

```{r}
#merging the dataframes by year_month
oil_merged <- merge(xom_monthly, oil_monthly, by = "year_month")

#plotting with xon as y-variabel and oil as x-variabel
oil_merged %>%
  ggplot(aes(x=exxon_avg, y=oil_avg)) +
  geom_smooth(method=lm, se = FALSE)+
  geom_abline(slope=1, intercept = 0, color="red") +
  labs(title="Correlation between XOM
and oil price", subtitle='The blue line is the actual correlation', y="Oil", x="XOM") +
  theme_bw()


```

The blue line shows the correlation between exxon and brent oil.

```{r}
#Use R’s  lm() function. Set the variable on the y-axis and x-axis, and specify the data set. 
lm(oil_avg ~ exxon_avg, data=oil_merged)
```
